Language-Based Security for Low-Level MPC

### 2024-06-25 revision-2/example-bdoz.tex
checkpoint

We also define a function $\ttt{secopen}$ to implement ``secure
opening''.  In this standard subprotocol, the value
$\macgv{\secret{w_1}} \fplus \macgv{\mesg{w_2}}$ is reconstructed as
$\mesg{w_3}$, by each client $\cid_2$ communicating
$\lcod{\mesg{w_1} + \mesg{w_2}}{\cid_2}$ to $\cid_1$.  Assuming that
$\macgv{\mesg{w_2}}$ is in an independent uniform distribution,
this perfectly hides $\secret{w_1}$. In a mutiplication gate
either $a$ or $b$ of a Beaver triple are used in secure opening,
so, e.g., given $
(\varnothing,\ttt{secopen}(a, x, d, 2, 1) \redxs (\prog,())$
we have both:
\begin{mathpar}
  \conddetx{\progtt(\prog)}{\{\macgv{\mesg{a}},\elab{\secret{x}}{1} \}}{\{ \elab{\mesg{d}}{2} \}}

@@

The full protocol for malicious secure product of secrets $x$ (that
is, $\elab{\secret{x}}{1}$) and $y$ (that is, $\elab{\secret{y}}{2}$)
using Beaver triple $a,b,c$ is defined in Figure \ref{fig-bdoz}. Both
parties interact in secure opening of $x \fplus a$ and $y + b$,
followed by the non-interactive computation of shares of $x * y$ for
secure reveals as per standard protocol
\cite{10.1007/978-3-030-68869-1_3}. The non-interactive reconstruction
of the local authentication key for both the secure openings and the
final reveal is enabled by the semi-homorphic properties of the BDOZ
scheme.

@@

We can carry out similar proofs of passive security for the protocol
in Figure \ref{fig-beaver} as for GMW, even using automated tactics
for the protocol in $\mathbb{F}_2$. But in the case of BDOZ we are
also concerned with malicious security. To demonstrate this, we can
show that the protocol satisfies integrity in the sense of Definition
\ref{def-integrity}. To do so, we observe that it satisfies a stronger
property, that we call cheating detection. Cheating detection says
that the adversary can only execute the protocol honestly, or or else
gets caught (and abort). Note in particular that this is \emph{not}
a form of endorsement (the dual of declassificatio), which allows
the adversary some leeway. Rather, it is a confirmation of high integrity.

>>>>

We also define a function $\ttt{secopen}$ to implement ``secure
opening''.  In this standard subprotocol, the value
$\macgv{\secret{w_1}} \fplus \macgv{\mesg{w_2}}$ is reconstructed as
$\mesg{w_3}$, by each client $\cid_2$ communicating
$\lcod{\mesg{w_1} + \mesg{w_2}}{\cid_2}$ to $\cid_1$.  Assuming that
$\macgv{\mesg{w_2}}$ is in an independent uniform distribution,
this perfectly hides $\secret{w_1}$. In a multiplication gate
either $a$ or $b$ of a Beaver triple are used in secure opening,
so, e.g., given $
(\varnothing,\ttt{secopen}(a, x, d, 2, 1) \redxs (\prog,())$
we have both:
\begin{mathpar}
  \conddetx{\progtt(\prog)}{\{\macgv{\mesg{a}},\elab{\secret{x}}{1} \}}{\{ \elab{\mesg{d}}{2} \}}

@@

The full protocol for malicious secure product of secrets $x$ (that
is, $\elab{\secret{x}}{1}$) and $y$ (that is, $\elab{\secret{y}}{2}$)
using Beaver triple $a,b,c$ is defined in Figure \ref{fig-bdoz}. Both
parties interact in secure opening of $x \fplus a$ and $y + b$,
followed by the non-interactive computation of shares of $x * y$ for
secure reveals as per standard protocol
\cite{10.1007/978-3-030-68869-1_3}. The non-interactive reconstruction
of the local authentication key for both the secure openings and the
final reveal is enabled by the semi-homomorphic properties of the BDOZ
scheme.

@@

We can carry out similar proofs of passive security for the protocol
in Figure \ref{fig-beaver} as for GMW, even using automated tactics
for the protocol in $\mathbb{F}_2$. But in the case of BDOZ we are
also concerned with malicious security. To demonstrate this, we can
show that the protocol satisfies integrity in the sense of Definition
\ref{def-integrity}. To do so, we observe that it satisfies a stronger
property, that we call cheating detection. Cheating detection says
that the adversary can only execute the protocol honestly, or or else
gets caught (and abort). Note in particular that this is \emph{not}
a form of endorsement (the dual of declassification), which allows
the adversary some leeway. Rather, it is a confirmation of high integrity.

### 2024-06-25 revision-2/examples.tex
checkpoint

In a malicious setting, detecting cheating by adding
information-theoretic secure MAC codes to shares is a fundamental approach
realized by protocols such as BDOZ \cite{XXX} and SPDZ \cite{XXX}.
These protocols assume a pre-processing phase that distributes shares
with MAC codes to clients.  This integrates well with pre-processed
Beaver Triples to implement malicious secure, and relatively
efficient, multiplication \cite{XXX}.  Here we show how to implement
malicious secure two party multiplication in the BDOZ style in
$\metaprot$.

A field value $v$ is secret shared among 2 client in BDOZ in
the same manner as in GMW.
Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where
$v_\cid$ is a secret share of $v$, i.e., $v = \fcod{v_1 \fplus v_2}$.
Further, field values $m_\cid$ are MACs of $v_\cid$ that are authenticated
by the other client's key. In particular, $m_\cid = k + (k_\Delta *
v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known only to $\cid'
\ne \cid$ and $k_\Delta$. Further, there is a unique \emph{local key}
$k$ for each shared value, while the \emph{global key} $k_\Delta$ is
common to all MACs authenticated by $\cid'$. By extending certain
field operations to these MACed values, a semi-homorphic encryption
scheme is obtained that is adequate to compute Beaver Triples. For
example, if local keys $k_1$ and $k_2$ authenticate $\macv_1$ and
$\macv_2$, then $k_1 \fplus k_2$ authenticates $\macv_1 \macplus
\macv_2$, where $\macplus$ is addition extended to MACed
values. Semantics for summation of MACed shares, and multipication of
shares and plain field values, are given in Figure \ref{fig-bdoz}.
For more details the reader is referred to \cite{XXX}. We note
that while we restrict values $v$, $m$, and $k$ to the same
field in this presentation for simplicity, in general $m$ and
$k$ can be in extensions of $\mathbb{Z}_p$.

We define variable forms $\macx{\secret{w}}{\cid}$ and
$\macx{\flip{w}}{\cid}$ to range over local MACed shares of a secret
and random sample owned by client $\cid$, respectively. We let
$\mack{x}{}$ denote the local key for authenticating shares of $x$.
We also assume a global key $\flip{\ttt{delta}}$ on each client.
Pre-processing is modeled by assuming that shares are provided on
input random tapes (memories) with the required properties, as defined
on the bottom of Figure \ref{fig-bdoz}.

Similarly to field values, we can use MACed shares in
uniform independent distributions as one-time-pads for
hiding other values, using subtraction over MACed shares.
So we introduce a function $\macotp$ to do this
declaratively.
$$
\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
$$

@@

\begin{definition}[BDOZ preprocessing]
  Define:
  \begin{mathpar}
    \mathit{shares} \defeq
    \{ \elab{\mesg{w}}{\cid} | \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid} | \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid} | \cid \in \{ 1, 2 \} \wedge w \in
    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid} | \cid \in \{ 1, 2 \} \} \end{array}
  \end{mathpar}
  Then a memory $\store$ satisfies BDOZ preprocessing iff:
  $$\dom(\store) = \mathit{shares}  \cup \mathit{macs} \cup \mathit{keys}$$
  and, writing $\store(\macgv{\mesg{w}})$ to denote
  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
  the following conditions hold for all $\cid \in \{1,2\}$ and $w \in \{ a,b,c,x,y\}$:
\begin{mathpar}
  \store(\macgv{\mesg{w@\cid}}) = \store(\elab{\secret{w}}{\cid})

  \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})

  \lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}  \qquad \cid \ne \cid'  
\end{mathpar}
\end{definition}

>>>>

In a malicious setting, detecting cheating by adding
information-theoretic secure MAC codes to shares is a fundamental
approach realized by protocols such as BDOZ \cite{XXX} and SPDZ
\cite{XXX}.  These protocols assume a pre-processing phase that
distributes shares with MAC codes to clients.  This integrates well
with pre-processed Beaver Triples to implement malicious secure, and
relatively efficient, multiplication \cite{XXX}. Recall
that Beaver triples are values $a,b,c$ with $\fcod{a * b = c}$,
unique per multiplication gate, that are secret shared with clients
during pre-processing.

A field value $v$ is secret shared among 2 client in BDOZ in the same
manner as in GMW.  Each client $\cid$ gets a pair of the form
$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
\emph{global key} $k_\Delta$ is common to all MACs authenticated by
$\cid'$. This is a semi-homomorphic encryption scheme that supports
addition of shares and multiplication of shares by a constant
\cite{XXX}. For more details the reader is referred to \cite{XXX}. We
note that while we restrict values $v$, $m$, and $k$ to the same field
in this presentation for simplicity, in general $m$ and $k$ can be in
extensions of $\mathbb{Z}_p$.

We can capture both the distribution of Beaver triples and
BDOZ shares as a pre-processing predicate that establishes
conditions for initial memories (see Definition \ref{def-progttb}).
As for GMW, we use $\macgv{\mesg{w}}$ to refer to secret-shared
values reconstructed with addition, where by convention
shares are message values $\elab{\mesg{w}}{\cid}$ for all $\cid$.
\begin{definition}[BDOZ preprocessing]
  Define:
  \begin{mathpar}
    \mathit{shares} \defeq
    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in
    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \} \end{array}
  \end{mathpar}
  Then a memory $\store$ satisfies BDOZ preprocessing iff:
  $$\dom(\store) = \{ \elab{\secret{x}}{1}, \elab{\secret{y}}{2} \} \cup \mathit{shares}
  \cup \mathit{macs} \cup \mathit{keys}$$
  and, writing $\store(\macgv{\mesg{w}})$ to denote
  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
  the following conditions hold:
  \begin{mathpar}
    \store(\macgv{\mesg{x}}) = \store(\elab{\secret{x}}{1})

    \store(\macgv{\mesg{y}}) = \store(\elab{\secret{y}}{2})

    \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
  \end{mathpar}
  and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
\end{definition}

\input{beaver}

@@

\begin{definition}[BDOZ preprocessing]
  Define:
  \begin{mathpar}
    \mathit{shares} \defeq
    \{ \elab{\mesg{w}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{macs} \defeq  \{ \elab{\mesg{w\ttt{mac}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in \{ a,b,c,x,y \}  \}

    \mathit{keys} \defeq  \begin{array}{l}\{ \elab{\mesg{w\ttt{k}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \wedge w \in
    \{ a,b,c,x,y \}  \} \cup \\ \{ \elab{\mesg{\ttt{delta}}}{\cid}\ |\ \cid \in \{ 1, 2 \} \} \end{array}
  \end{mathpar}
  Then a memory $\store$ satisfies BDOZ preprocessing iff:
  $$\dom(\store) = \{ \elab{\secret{x}}{1}, \elab{\secret{y}}{2} \} \cup \mathit{shares}
  \cup \mathit{macs} \cup \mathit{keys}$$
  and, writing $\store(\macgv{\mesg{w}})$ to denote
  $\fcod{\store(\elab{\mesg{w}}{1}) + \store(\elab{\mesg{w}}{2})}$,
  the following conditions hold:
  \begin{mathpar}
    \store(\macgv{\mesg{x}}) = \store(\elab{\secret{x}}{1})

    \store(\macgv{\mesg{y}}) = \store(\elab{\secret{y}}{2})

    \fcod{\store(\macgv{\mesg{a}}) * \store(\macgv{\mesg{b}})} = \store(\macgv{\mesg{c}})
  \end{mathpar}
  and for all $\cid,\cid' \in \{1,2\}$ with $\cid \ne \cid'$ and $w \in \{ a,b,c,x,y\}$:
  $$\lcod{\store, \mesg{w\ttt{mac}}}{\cid} =
  \lcod{\store, \mesg{wk} + \mesg{\ttt{delta}} * \mesg{w}}{\cid'}$$
\end{definition}

### 2024-06-25 revision-2/lang.tex
checkpoint

The $\minifed$ language provides a simple model of synchronous
protocols between a federation of \emph{clients} exchanging values in
the binary field. We will identify clients by natural numbers, and
federations- finite sets of clients- are always given statically.
As we will see, our threat model assumes a partition of the federation
into \emph{honest} $H$ and \emph{corrupt} $C$ subsets.

We model probabilistic programming via a \emph{random tape}
semantics. That is, we will assume that programs can make reference to
values chosen from a uniform random distributions defined in the
initial program memory.  Programs aka protocols execute
deterministically given the random tape.

@@

The syntax of $\minifed$, defined in
Figure \ref{fig-minifed}, includes values $v$ and standard
operations of addition, subtraction, and multiplication in
a finite field $\mathbb{F}_p$ where $p$ is some prime. 
Protocols are given input secret values $\secret{w}$
as well as random samples $\flip{w}$ on the input
tape, both of which are distinguished by
strings $w$. Protocols are sequences of assignment
commands of three different forms:
\begin{itemize}
\item $\eassign{\mesg{w}}{\cid_2}{\be}{\cid_1}$: This
  is a \emph{message send} where expression $\be$ is computed
  by client $\cid_1$ and sent to client $\cid_2$ as message
  $\mesg{w}$.
\item $\reveal{w}{\be}{\cid}$: This
  is a \emph{public reveal} where expression $\be$ is computed
  by client $\cid$ and broadcast to the federation, typically
  to communicate intermediate results for use in final output
  computations.
\item $\pubout{\cid}{\be}{\cid}$: This
  is an \emph{output} where expression $\be$ is computed
  by client $\cid$ and reported as its output. As a
  sanity condition we disallow commands
  $\pubout{\cid_1}{\be}{\cid_2}$ where $\cid_1\ne\cid_2$.
\end{itemize}

Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be
referenced in expressions, once they've been assigned.
This distinction between messages and broadcast public
reveal is consistent with previous formulations, e.g.,
\cite{6266151}.

We let $x$ range over \emph{variables} which are identifiers where
client ownership is specified- e.g.,
$\elab{\mesg{\mathit{foo}}}{\cid}$ is a message $\mathit{foo}$ that
was sent to $\cid$. We let $X$ range over sets of variables, and more
specifically, $S$ ranges over sets of secret variables
$\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables
$\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables
$\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables
$\rvl{w}$, and $O$ ranges over sets of output variables $\out{\cid}$.
Given a program $\prog$, we write $\iov(\prog)$ to denote the set $S
\cup M \cup P \cup O$ of variables in $\prog$ with ownership made
explicit and $\secrets(\prog)$ to denote $S$, and we write
$\flips(\prog)$ to denote the set $R$ of random samplings in $\prog$
with ownership made explicit. We write $\vars(\prog)$ to denote
$\iov(\prog) \cup \flips(\prog)$. For any set of variables $X$ and
parties $P$, we write $X_P$ to denote the subset of $X$ owned by any
party in $P$, in particular we write $X_H$ and $X_C$ to denote the
subsets belonging to honest and corrupt parties, respectively.

\subsection{Semantics}

@@

Given a variable-free expression $\be$, we write $\cod{\be}$ to denote
the standard interpretation of $\be$ in the arithmetic field
$\mathbb{Z}_{p}$. With the introduction of variables to expressions,
we need to interpret variables with respect to a specific memory, and
all variables used in an expression must belong to a specified client.
Thus, we denote interpretation of expressions $\be$ computed on a
client $\cid$ as $\lcod{\store,\be}{\cid}$. This interpretation is
defined in Figure \ref{fig-minifed}. It is also parameterized by
$\delta$ which defines the semantics of library functions $f$.

The small-step reduction relation $\redx$ is then defined in Figure
\ref{fig-minifed} to evaluate commands. Reduction is a relation on
\emph{configurations} $(\store, \prog)$ where all three command forms-
message send, broadcast, and output- are implemented as updates to the
memory $\store$. We write $\redxs$ to denote the reflexive, transitive
closure of\ $\redx$.

@@

Shamir addition leverages homomorphic properties of addition in
arithmetic fields to implement secret addition. If a field value $v_1$
is in a uniform random distribution with other variables in a program,
then either  $v_1 \fplus v_2$ or $v_1 \fminus v_2$  is an encryption of $v_2$
where $v_1$ is an information theoretically secure one-time-pad, which
is exploited for secret sharing. Of course, $\fplus$ and $\fminus$ are also
meaningful operations over any two field values regardless of their distributions.

To sum their secret values $\secret{\cid}$, each client $\cid$ in
the federation $\{ 1, 2, 3 \}$  samples a value $\locflip$
that can be used as a one-time pad with another
random sample $\flip{x}$ and $\secret{\cid}$. This yields
two secret shares communicated as messages to the other clients,
while each client keeps $\locflip$ as its own share.
$$
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{1})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{2})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\locflip \fminus \flip{x} \fminus \secret{3})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array}
$$
Due to field properties of $\fminus$ this scheme guarantees that messages
are viewed as random noise by any observer 
besides $\cid$ \cite{barthe2019probabilistic}. Next, each client
publicly reveals the sum of all of its shares, including its local
share. This step does reveal information about secrets-- note in
particular that $\locflip$ is reused and is no longer a one-time-pad:
$$
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
\end{array}
$$
Finally, each client outputs the sum of each sum of shares, yielding
the sum of secrets. The protocol is correct because the outputs are all the
true sum of secrets, and it is secure because no more information about the
secrets other than that revealed by their sum is exposed.
$$
%\elab{\mesg{o1}}{2} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  %\elab{\mesg{o1}}{3} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  %\elab{\mesg{o2}}{1} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  %\elab{\mesg{o2}}{3} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\ 
  %\elab{\mesg{o3}}{1} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\ 
  %\elab{\mesg{o3}}{2} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3}\\ 
  %\pubout{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3} + \mesg{o2} + \mesg{o3})}{1}
\begin{array}{lll}
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
$$

>>>>

The $\minifed$ language establishes a basic model of synchronous
protocols between a federation of \emph{clients} exchanging values in
the binary field. A model of synchronous communication captures a wide
range of MPC protocols. Concurrency is out of scope in this work but
an avenue for future work. The lack of sophisticated control
structures in $\minifed$ is intentional, since minimizing features
eases analysis and control abstractions such as function definitions
can be integrated into a metalanguage that generates $\minifed$
programs (Section \ref{section-metalang}).

We identify clients by natural numbers and federations- finite sets of
clients- are always given statically.  Our threat model assumes a
partition of the federation into \emph{honest} $H$ and \emph{corrupt}
$C$ subsets. We model probabilistic programming via a \emph{random
tape} semantics. That is, we will assume that programs can make
reference to values chosen from a uniform random distributions defined
in the initial program memory.  Programs aka protocols execute
deterministically given the random tape.

@@

The syntax of $\minifed$, defined in Figure \ref{fig-minifed},
includes values $v$ and standard operations of addition, subtraction,
and multiplication in a finite field $\mathbb{F}_p$ where $p$ is some
prime.  Protocols are given input secret values $\secret{w}$ as well
as random samples $\flip{w}$ on the input tape, implemented using a
\emph{memory} as described below (Section
\ref{section-lang-semantics}) where $w$ is a distinguishing 
identifier string. Protocols are sequences of assignment commands of three
different forms:
\begin{itemize}
\item $\eassign{\mesg{w}}{\cid_2}{\be}{\cid_1}$: This
  is a \emph{message send} where expression $\be$ is computed
  by client $\cid_1$ and sent to client $\cid_2$ as message
  $\mesg{w}$.
\item $\reveal{w}{\be}{\cid}$: This
  is a \emph{public reveal} where expression $\be$ is computed
  by client $\cid$ and broadcast to the federation, typically
  to communicate intermediate results for use in final output
  computations.
\item $\pubout{\cid}{\be}{\cid}$: This
  is an \emph{output} where expression $\be$ is computed
  by client $\cid$ and reported as its output. As a
  sanity condition we disallow commands
  $\pubout{\cid_1}{\be}{\cid_2}$ where $\cid_1\ne\cid_2$.
\end{itemize}
For example, in the following protocol, a client 1
subtracts a random sample $\flip{y}$ from $\mathbb{F}_p$ from their
secret value $\secret{x}$ and sends the result to client
2 as a message $\mesg{z}$:
$$
\eassign{\mesg{z}}{2}{(\secret{x} - \flip{y})}{1}
$$
Both messages $\mesg{w}$ and reveals $\rvl{w}$ can be
referenced in expressions once they've been defined.
This distinction between messages and broadcast public
reveal is consistent with previous formulations, e.g.,
\cite{6266151}.

We let $x$ range over \emph{variables} which are identifiers where
client ownership is specified- e.g.,
$\elab{\mesg{\mathit{foo}}}{\cid}$ is a message $\mathit{foo}$ that
was sent to $\cid$. We let $X$ range over sets of variables, and more
specifically, $S$ ranges over sets of secret variables
$\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables
$\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables
$\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables
$\rvl{w}$, and $O$ ranges over sets of output variables $\out{\cid}$.
Given a program $\prog$, we write $\iov(\prog)$ to denote the set $S
\cup M \cup P \cup O$ of variables in $\prog$ with ownership made
explicit and $\secrets(\prog)$ to denote $S$, and we write
$\flips(\prog)$ to denote the set $R$ of random samplings in $\prog$
with ownership made explicit. We write $\vars(\prog)$ to denote
$\iov(\prog) \cup \flips(\prog)$. For any set of variables $X$ and
clients $I$, we write $X_I$ to denote the subset of $X$ owned by any
client $\cid \in I$, in particular we write $X_H$ and $X_C$ to denote the
subsets belonging to honest and corrupt parties, respectively.

\subsection{Semantics}
\label{section-lang-semantics}

@@

Given a variable-free expression $\be$, we write $\cod{\be}$ to denote
the standard interpretation of $\be$ in the arithmetic field
$\mathbb{Z}_{p}$. With the introduction of variables to expressions,
we need to interpret variables with respect to a specific memory, and
all variables used in an expression must belong to a specified client.
Thus, we denote interpretation of expressions $\be$ computed on a
client $\cid$ as $\lcod{\store,\be}{\cid}$. This interpretation is
defined in Figure \ref{fig-minifed}. The small-step reduction relation
$\redx$ is then defined in Figure \ref{fig-minifed} to evaluate
commands. Reduction is a relation on \emph{configurations} $(\store,
\prog)$ where all three command forms- message send, broadcast, and
output- are implemented as updates to the memory $\store$. We write
$\redxs$ to denote the reflexive, transitive closure of\ $\redx$.

@@

Shamir addition leverages homomorphic properties of addition in
arithmetic fields to implement secret addition. If a field value $v_1$
is uniformly random, then $v_1 \fminus v_2$ is an encryption of $v_2$
where $v_1$ is an information theoretically secure one-time-pad, which
is exploited for secret sharing, noting that $v_2$ can be
reconstructed by summing $v_1$ and $v_3 \defeq v_1 \fminus v_2$.

In $\minifed$, to privately sum secret values $\secret{\cid}$, each
client $\cid$ in the federation $\{ 1, 2, 3 \}$ samples a value
$\locflip$ that can be used as a one-time pad with another random
sample $\flip{x}$ and $\secret{\cid}$. This yields two secret shares
communicated as messages to the other clients, while each client keeps
$\locflip$ as its own share.
$$
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array}
$$
This scheme guarantees that messages
are viewed as random noise by any observer 
besides $\cid$ \cite{barthe2019probabilistic}. Next, each client
publicly reveals the sum of all of its shares, including its local
share. This step does reveal information about secrets-- note in
particular that $\locflip$ is reused and is no longer a one-time-pad:
$$
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
\end{array}
$$
Finally, each client outputs the sum of each sum of shares, yielding
the sum of secrets. The protocol is correct because the outputs are all the
true sum of secrets, and it is secure because no more information about the
secrets other than that revealed by their sum is exposed.
$$
%\elab{\mesg{o1}}{2} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  %\elab{\mesg{o1}}{3} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  %\elab{\mesg{o2}}{1} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  %\elab{\mesg{o2}}{3} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\ 
  %\elab{\mesg{o3}}{1} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\ 
  %\elab{\mesg{o3}}{2} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3}\\ 
  %\pubout{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3} + \mesg{o2} + \mesg{o3})}{1}
\begin{array}{lll}
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
$$

### 2024-06-25 revision-2/metalang.tex
checkpoint

The $\fedprot$ language is low-level and does not include abstractions
for defining composable elements. So in this Section we introduce the
$\metaprot$ language which that includes structured data and function
definitions for defining composable protocol elements at a higher
level of abstraction.  The $\metaprot$ language is a
\emph{metalanguage}, where $\fedprot$ protocols are the residuum of
computation. In addition to these declarative benefits of $\metaprot$,
component definitions support compositional verification of larger
prortocols as we will discuss with examples in Sections
\ref{section-example-gmw} and \ref{section-example-bdoz}.

@@

We define a small-step evaluation aka reduction relation $\redx$ in
Figure \ref{fig-metaprot}.  We write $\redxs$ to denote the
reflexive, transitive closure of $\redx$. Reduction is defined on
\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
where $\prog$ is the $\minifed$ program accumulated during evaluation.
In this definition we write $e[\mv/y]$ to denote the substitution of $\mv$
for free occurences of $y$ in $e$. The rules are mostly standard,
except when a concrete $\minifed$ assignment is encountered it is added
to the end of $\prog$.

>>>>

The $\fedprot$ language is low-level and does not include abstractions
for defining composable elements. So in this Section we introduce the
$\metaprot$ language which that includes structured data and function
definitions for defining composable protocol elements at a higher
level of abstraction.  The $\metaprot$ language is a
\emph{metalanguage}, where $\fedprot$ protocols are the residuum of
computation. In addition to these declarative benefits of $\metaprot$,
component definitions support compositional verification of larger
protocols as we will discuss with examples in Sections
\ref{section-example-gmw} and \ref{section-example-bdoz}.

@@

We define a small-step evaluation aka reduction relation $\redx$ in
Figure \ref{fig-metaprot}.  We write $\redxs$ to denote the
reflexive, transitive closure of $\redx$. Reduction is defined on
\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
where $\prog$ is the $\minifed$ program accumulated during evaluation.
In this definition we write $e[\mv/y]$ to denote the substitution of $\mv$
for free occurrences of $y$ in $e$. The rules are mostly standard,
except when a concrete $\minifed$ assignment is encountered it is added
to the end of $\prog$.
